PANCHANGE_descriptives

### 2020-11-03 PANCHANGE_exclusion_descriptives.Rmd
N sign up per month, impute retrospective if no change, retro and prepan correlations and drop na from education comparison tables

Load packages
```{r Load packages}
library(knitr)
library(summarytools)
library(psych)
library(polycor)
library(corrplot)
library(patchwork)
library(broom)
library(tidyverse)
library(ggpubr)
```

@@

## Read in data
```{r read in data}
dat.raw <- readRDS(file = paste0(data_path, "/four_cohorts_variables.rds"))
dim(dat.raw)
colnames(dat.raw)
```

@@

Exclude GLAD/EDGI individuals that didn't take part in COPING
Exclude individuals with NA for the MHD categories
Individuals with NAs for age/gender have already been excluded in pre-processing script, but they get excluded here to be save and to double check previous programming.

@@

@@

@@

# highest_education
## Proportion tables
```{r highest_education per Sample prop per total}
highest_education_per_Sample_prop_total <- dat %>%
  count(highest_education, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

@@

```{r highest_education per Sample prop per Sample}
highest_education_per_Sample_prop_Sample <- dat %>%
  group_by(Sample) %>%
  count(highest_education, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

@@

## Cross tables
```{r highest_education per Sample prop cross table}
highest_education_per_Sample_prop_Sample_cross_table <- dat %>%
  group_by(Sample) %>%
  count(highest_education, Sample) %>%
  mutate(Prop = round(n/sum(n), 2)) %>%
  pivot_wider(names_from = Sample, values_from = c(n, Prop))

@@

```{r highest_education per Sample count cross table}
highest_education_per_Sample_count_cross_table <- dat %>%
  count(highest_education, Sample) %>%
  spread(key = Sample, value = n) %>%
  mutate(
    EDGI = as.integer(replace_na(EDGI, 0)),
    GLAD = as.integer(replace_na(GLAD, 0)),
    NBR = as.integer(replace_na(NBR, 0)),
    RAMP = as.integer(replace_na(RAMP, 0))
  )

>>>>

Load packages
```{r Load packages}
library(knitr)
library(summarytools)
library(psych)
library(polycor)
library(corrplot)
library(patchwork)
library(broom)
library(tidyverse)
library(ggpubr)
library(stringr)
```

@@

## Read in data
```{r read in data}
dat.raw <- readRDS(file = paste0(data_path, "four_cohorts_variables.rds"))
dim(dat.raw)
colnames(dat.raw)
```

@@

Exclude GLAD/EDGI individuals that didn't take part in COPING
Exclude individuals with NA for the MHD categories
Individuals with NAs for age/gender have already been excluded in pre-processing script, but they get excluded here to be safe and to double check previous programming.

@@

#SB - Impute no change prepandemic scores

I interpreted the presence of an NA in the retrospective score as an indicator of saying "no change" in their baseline scores. Hence, if this was the case, I imputed their baseline score. However, this interpretation could be wrong.

Also, I'm not sure what to do if there is an NA for their baseline sum score (nothing to duplicate). This was the case for ocir for some reason: the retro scores were provided for some by not the corresponding baseline score?

```{r Impute for gad measure}

#Visualise the gad summary scores for baseline and retro
head(dat$gad.sum_score_base, 30)
head(dat$gad.sum_score_retro, 30)

#Edit the gad sum retro score as follows: if there is a missing value in the retro sum score, give it the baseline sum score, else leave it as it is

impute_dat <- dat %>% 
  mutate(gad.sum_score_retro = 
           if_else(is.na(dat$gad.sum_score_retro), dat$gad.sum_score_base, dat$gad.sum_score_retro))

#Check that the Na's have been replaced with duplicated baseline sum scores
head(impute_dat$gad.sum_score_base, 30)
head(impute_dat$gad.sum_score_retro, 30)

```{r Impute for phq measure}

#Visualise the phq summary scores for baseline and retro
head(dat$phq.sum_score_base, 30)
head(dat$phq.sum_score_retro, 30)

#Edit the phq sum retro score as follows: if there is a missing value in the retro sum score, give it the baseline sum score, else leave it as it is

impute_dat <- dat %>% 
  mutate(phq.sum_score_retro = 
           if_else(is.na(dat$phq.sum_score_retro), dat$phq.sum_score_base, dat$phq.sum_score_retro))

#Check that the Na's have been replaced with duplicated baseline sum scores
head(impute_dat$phq.sum_score_base, 30)
head(impute_dat$phq.sum_score_retro, 30)

```{r Impute for ocir measure}

#Visualise the ocir summary scores for baseline and retro
head(dat$ocir.sum_score_base, 40)
head(dat$ocir.sum_score_retro, 40)

#Edit the ocir sum retro score as follows: if there is a missing value in the retro sum score, give it the baseline sum score, else leave it as it is

impute_dat <- dat %>% 
  mutate(ocir.sum_score_retro = 
           if_else(is.na(dat$ocir.sum_score_retro), dat$ocir.sum_score_base,                                     dat$ocir.sum_score_retro))

#Check that the Na's have been replaced with duplicated baseline sum scores
head(impute_dat$ocir.sum_score_base, 40)
head(impute_dat$ocir.sum_score_retro, 40)

```{r Impute for pcl measure}

#Visualise the pcl summary scores for baseline and retro
head(dat$pcl.sum_score_base, 30)
head(dat$pcl.sum_score_retro, 30)

#Edit the pcl sum retro score as follows: if there is a missing value in the retro sum score, give it the baseline sum score, else leave it as it is

impute_dat <- dat %>% 
  mutate(pcl.sum_score_retro = 
           if_else(is.na(dat$pcl.sum_score_retro), dat$pcl.sum_score_base, dat$pcl.sum_score_retro))

#Check that the Na's have been replaced with duplicated baseline sum scores
head(impute_dat$pcl.sum_score_base, 30)
head(impute_dat$pcl.sum_score_retro, 30)

#SB - Baseline survey completion (N per month)

This is a summary of the amount of pt's completing the baseline survey per month

```{r N completing coping per month}

#Visualise some dates when participants completed the questionnaire
head(dat$startDate.coping, 20)

#Divide the data by sign up per month and summarise the amount of pt's completing the survey that month

april_sign_up <- dat[str_detect(dat$startDate.coping, "-04-"), ] %>% 
  select(startDate.coping) %>% 
  summarise(April = n())

may_sign_up <- dat[str_detect(dat$startDate.coping, "-05-"), ] %>% 
  select(startDate.coping) %>% 
  summarise(May = n())

june_sign_up <- dat[str_detect(dat$startDate.coping, "-06-"), ] %>% 
  select(startDate.coping) %>% 
  summarise(June = n())

july_sign_up <- dat[str_detect(dat$startDate.coping, "-07-"), ] %>% 
  select(startDate.coping) %>% 
  summarise(July = n())

august_sign_up <- dat[str_detect(dat$startDate.coping, "-08-"), ] %>% 
  select(startDate.coping) %>% 
  summarise(August = n())

september_sign_up <- dat[str_detect(dat$startDate.coping, "-09-"), ] %>% 
  select(startDate.coping) %>% 
  summarise(September = n())

#Merge results into one tibble

coping_sign_up <- april_sign_up %>% 
  cbind(may_sign_up, june_sign_up, july_sign_up, august_sign_up, september_sign_up)

#Gather the variables
coping_sign_up <- coping_sign_up %>% 
  gather(key = "month", value = "sign_up_number")

#Proportion calculation
coping_sign_up <- coping_sign_up %>% 
  mutate(percentage = 
           round((sign_up_number / sum(sign_up_number) * 100), 2))

coping_sign_up

@@

#SB - Correlation between retropsective impute and prepan

```{r Correlation prep chunk}
# Select items for correlation matrix
prepan_retro_items <- c(
    "gad.sum_score_prepan",
    "gad.sum_score_retro", 
    "pcl.sum_score_prepan",
    "pcl.sum_score_retro",
    "ocir.sum_score_prepan",
    "ocir.sum_score_retro",
    "phq.sum_score_prepan",
    "phq.sum_score_retro"
)

#Create a data set with only the COPING samples
prepan_retro_dat <- dat[str_detect(dat$Sample, "NBR|GLAD|EDGI"), ]

#Check that RAMP has been removed
prepan_retro_dat %>% 
  freq(Sample)

# Calculate correlation matrix on specific columns (variables) from your data frame
prepan_retro.corr.mat <- hetcor(
  as.data.frame(prepan_retro_dat[,prepan_retro_items]), # Specify the data
  use = "pairwise.complete.obs" # Use all pairwise observations
)

# Print the correlation matrix
print(prepan_retro.corr.mat, digits = max(3, getOption("digits") - 3))

# Save the correlations from the correlation matrix in an object
prepan_retro.corr.matrix <- as.matrix(prepan_retro.corr.mat)

# Save the p values from the correlation matrix in an object
prepan_retro.p.matrix <- as.matrix(prepan_retro.corr.mat$tests)
```

```{r Correlation matrix of prepan base sum scores}
corrplot(prepan_retro.corr.matrix, 
         method = "color", # objects to represent the correlations on plot
         type = "lower", # only use the lower triangle of the matrix
         diag = FALSE, # do not show the correlations on the diagonal
         addgrid.col = NA,
         addCoef.col = "black", # colour for the correlation coefficients in the plot
         tl.cex = 0.8,
         tl.col = "black",
         col=colorRampPalette(c("dodgerblue4","white","firebrick4"))(200), # colours for correlations
         # Combine with significance
         p.mat = prepan_retro.p.matrix, # Matrix with p values
         sig.level = 0.01, # Choose significant level
         insig = "blank" # Nonsignificant correlations have no colour
         )
```

@@

# highest_education
## Proportion tables
```{r highest_education per Sample prop per total}
highest_education_per_Sample_prop_total <- dat %>%
  count(highest_education, Sample) %>%
  drop_na() %>% 
  mutate(Prop = round(n/sum(n), 2))

@@

```{r highest_education per Sample prop per Sample}
highest_education_per_Sample_prop_Sample <- dat %>%
  group_by(Sample) %>%
  count(highest_education, Sample) %>%
  drop_na() %>% 
  mutate(Prop = round(n/sum(n), 2))

@@

## Cross tables
```{r highest_education per Sample prop cross table}
highest_education_per_Sample_prop_Sample_cross_table <- dat %>%
  group_by(Sample) %>%
  count(highest_education, Sample) %>%
  drop_na() %>% 
  mutate(Prop = round(n/sum(n), 2)) %>%
  pivot_wider(names_from = Sample, values_from = c(n, Prop))

@@

```{r highest_education per Sample count cross table}
highest_education_per_Sample_count_cross_table <- dat %>%
  count(highest_education, Sample) %>%
  drop_na() %>% 
  spread(key = Sample, value = n) %>%
  mutate(
    EDGI = as.integer(replace_na(EDGI, 0)),
    GLAD = as.integer(replace_na(GLAD, 0)),
    NBR = as.integer(replace_na(NBR, 0)),
    RAMP = as.integer(replace_na(RAMP, 0))
  )
